Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-01 finaletd/mainintroduction.tex
changes

Concurrent program is a term used to describe multiple threads / processes running simultaneously to perform some computation. There are two different environments where concurrent programs are used: shared memory and distributed memory systems. Shared memory systems contain multiple processors / multi-core processors sharing a common memory and execute the computation through threads / processes. Threads communicate using shared memory and may use atomic instructions or traditional locks to execute some critical sections of the program. 
In a shared memory environment, concurrent programmers use smart pointers to manually manage memory. Smart pointers are a manual reference counting technique implemented by the application programming interface. Distributed memory systems contain multiple processors with dedicated memory and processors are separated by some physical distance and connected via a network. In distributed memory systems, processes execute on each processor and  communicate by sending messages across the network to perform the computation. Both of the environments contain programs that run concurrently and require some form of communication to perform the desired computation. The environments are distinguished by communcation technique and memory use. Both of the environments uses automatic memory management feature to help developers build applications with above-mentioned benefits.

When concurrent programs are written by developers, the manual memory management is extremely difficult and error-prone. When multiple threads access a common object, the dangling pointers are common scenario due to the incomplete information about ownership. Double free bugs can occur just as frequent as the dangling pointers. Smart pointers are the defacto standard to avoid memory management issues in the concurrent programs. These smart pointers do not guarantee complete garbage collection unless until carefully used. The use of cyclic objects in the heap memory requires extreme care in a manually managed concurrent program. It is carefully avoided by all programmers usually to make the program memory leak free. The above reasons explain how difficult the automatic memory management in the concurrent programming environment is and the necessity for high-quality automatic memory management in concurrent programming environments.

>>>>

Concurrent program is a term used to describe multiple threads / processes running simultaneously to perform some computation. There are two different environments where concurrent programs are used: shared memory and distributed memory systems. Shared memory systems contain multiple processors / multi-core processors sharing a common memory and execute the computation through threads / processes. Threads communicate using shared memory and may use atomic instructions or traditional locks to execute some critical sections of the program. 
In a shared memory environment, concurrent programmers use smart pointers to manually manage memory. Smart pointers are a manual reference counting technique implemented by the application programming interface. Distributed memory systems contain multiple processors with dedicated memory and processors are separated by some physical distance and connected via a network. In distributed memory systems, processes execute on each processor and  communicate by sending messages across the network to perform the computation. Both of the environments contain programs that run concurrently and require some form of communication to perform the desired computation. The environments are distinguished by communcation technique and memory use. Both of the environments uses automatic memory management to help developers build applications with the above-mentioned benefits.

When concurrent programs are written by developers, the manual memory management is extremely difficult and error-prone. When multiple threads access a common object, dangling pointers are a common scenario due to the incomplete information about ownership. Double-free bugs can occur just as frequently as the dangling pointers. Smart pointers are the defacto standard to avoid memory management issues in concurrent programs. These smart pointers do not guarantee complete garbage collection. The use of cyclic objects in the heap memory requires extreme care in a manually managed concurrent program. It is carefully avoided by all programmers usually to make the program memory leak free. The above reasons explain how difficult the automatic memory management in the concurrent programming environment is and the necessity for high-quality automatic memory management in concurrent programming environments.
